\documentclass[11pt]{amsart}

\usepackage{amsthm, amssymb,amsmath}

\theoremstyle{definition}  % Heading is bold, text is roman
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}

\newcommand{\ojo}[1]{{\sffamily\bfseries\boldmath[#1]}}

\oddsidemargin 0pt
\evensidemargin 0pt
\marginparwidth 0pt
\marginparsep 10pt
\topmargin -10pt
\headsep 10pt
\textheight 8.4in
\textwidth 7in

%\input{header}


\begin{document}

%\exam{}{Take Home Exam I}

\noindent
This takehome exam is due 5:00 pm Friday, September 30. Your answers need to be typeset in \LaTeX \ and uploaded on iLearn. Please also read and sign the following and return it 
with your answers.

\vskip 3cm
\noindent While I may have consulted with Serkan regarding this exam, the solutions presented here are my own work. I have
not consulted any books other than the textbook and notes, nor  have I utilized the internet in any form in relation to the exam. I understand
that to get full credit, I need to show all the steps necessary to arrive at the answer, and unless it is obvious, explain 
my reasoning with complete sentences.


\vskip 1cm 
\noindent {\bf Name:}   \hspace{7cm}   {\bf Signature:}



\vfill
\eject

\noindent {\bf 1.a)} Let $\mathbb{Q}^+$ be the set of all nonnegative rational numbers. Show that $\mathbb{Q}^+$ is a group under addition. \\
{\bf 1.b)} Now let $S$ be the set of positive irrational numbers. Decide whether $S$ is a group under multiplication: either show that it is a group or provide
a reason why it is not a group. \\

\noindent {\bf 2.} Let $G$ be a group and let $a$ and $b$ be two elements in $G$. Suppose $n$ is a positive integer. Compute $(aba^{-1})^n$. [Induction !]\\


\noindent {\bf 3.} Let $G$ be a group and let $a$ and $b$ be any two elements in $G$. Suppose $n$ is a positive integer. Show that if $G$ is an abelian group then  $(ab)^n = a^nb^n$.
Give an example of a non-abelian group in which this equation does not always hold. \\


\noindent {\bf 4.a)} Let $G$ be a group and $g \in G$. Prove that $g^2 = e$ if and only if $g = g^{-1}$. \\
{\bf 4.b)} Show that if $g^2 = e$ for all $g \in G$ then $G$ is an abelian group. \\
  


\noindent {\bf 5.} Let $n > 2$ be an integer. Show that there exist at least two elements $[x]$ and $[y]$ in $U(n)$ such that $[x]^2 = [y]^2 = [1]$. 
\end{document}



